---
title: "Analysis of tooth growth in guinea pigs"
author: "Srijan Nayak"
date: "08/08/2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE, message = FALSE)
```

```{r load_libraries}
# load libraries for analysis
library(dplyr)
library(ggplot2)
```

```{r load_data}
# load the ToothGrowth data
data("ToothGrowth")
```

## Tooth length vs. supplement type and dose

Let's see how the response varies for different supplement types and dosages
with a boxplot.

```{r boxplot, fig.height=3.2}
ToothGrowth %>%
  mutate(
    dose = as.factor(dose),
    # orange juice is coded as 'OJ' and ascorbic acid is coded as 'VC'
    supp = if_else(supp == "OJ", "Orange juice", "Ascorbic acid")
  ) %>%
  ggplot() +
  geom_boxplot(aes(x = dose, y = len, color = supp)) +
  labs(title = "Effect of supplement type and dose on tooth length",
       x = "Dose", y = "Tooth length", color = "Supplement type")
```

From the above plot it seems that more dosage results in longer tooth length. It
also seems that orange juice might be a better supplement than ascorbic acid.

## Does the method of dosage matter?

The previous graph indicated that orange juice might better for tooth growth
than ascorbic acid. This can be tested with a t-test, where
$$
\mu_0: \bar{x}_{oj} - \bar{x}_{vc} = 0
$$
$$
\mu_a: \bar{x}_{oj} - \bar{x}_{vc} > 0
$$
Here $\bar{x}_{oj}$ is the mean tooth length for guinea pigs taking orange juice
as the vitamin C supplement while $\bar{x}_{vc}$ is the mean tooth length for
those taking ascorbic acid as the supplement.

```{r supplement_t_test}
tooth_length_oj <- filter(ToothGrowth, supp == "OJ") %>% select(len)
tooth_length_vc <- filter(ToothGrowth, supp == "VC") %>% select(len)
(method_p_value <- t.test(tooth_length_oj, tooth_length_vc)$p.value)
```

The test has a p-value of $`r round(method_p_value, 3)` > 0.05$, thus we fail to
reject the null hypothesis that **the method of dosage doesn't matter** at a
significance level of 5%.
